Crowdy: Social Crowd Visualization

### 2011-03-01 docs/crowdy_gui.tex
merged new related work, minor edits

\section{Related Work}
There has been some work into understanding how communities of users interact
on social networks. Along with working on understanding microblogging usage and
communities\cite{java2007we}, the main author - Akshay Java was one of the
first few who dealt with the measurement of the usage and nature of communities
in microblogging. In his latter study\cite{java2008mining}, he presented his
observations of the microblogging phenomena and user intentions by studying the
content, topological and geographical properties of such communities. He found
that microblogging provides users with a more immediate form of communication
to talk about their daily activities and to seek or share information.

When it comes to visualization of social crowds, a tool which has
similar objectives to ours is
Vizster\cite{heer2005vizster}. It is a visualization system for playful
end-user exploration and navigation of large-scale online social networks. The
design builds upon familiar node-link network layouts to contribute techniques
for exploring connectivity in large graph structures, supporting visual search
and analysis, and automatically identifying and visualizing community
structures. The system facilitates discovery of people, connections,
and communities to promote increased awareness of community structure and
information exposure, while maintaining a fun and socially engaged online
space. The creators of the tool were interested in crafting techniques for the
situated exploration and visual analysis of large graph structures by a
specific end-user population.

Another related study was on how web-based retrieval from Twitter can be used
as knowledge base\cite{cheong2009integrating} which shows that the
visualization system we try to attempt can be helpful. Marc Cheong et al. in
this study tried to analyze demographics of the authors of Twitter messages and
attempt to map a Twitter trend into what's going on in the real world. Their
findings reveal a pattern behind trends on Twitter, and trying to evolve
through visualization methods.  Their findings also enable us to understand the
underlying characteristics behind the ``trend setters'', providing us a new
perspective on the contributors of a trend.

Identifying highly dynamic ad-hoc collections of users what we refer to as
crowds in massive social messaging systems like Twitter and Facebook is
important\cite{kamath2010identifying}. K. Y. Kamath et al. in the study have a
salient feature of the proposed approach as an efficient locality-based
clustering approach for identifying crowds of users in near real-time compared
to more heavyweight static clustering algorithm. The study consisted of 711,612
users and 61.3 million messages, which show how the proposed approach can
efficiently and effectively identify Twitter-based crowds.

In the other study based on the previous one, K. Y. Kamath et
al\cite{kamath2011transient}, they add another salient feature - a novel crowd
tracking and evolution approach for linking crowds across time periods. Unlike
the more static and long-lived group-based membership offered on many social
networks their goal is to support the discovery of organic and highly-temporal
group affiliation, which they refer to as ``transient crowds''. A transient
crowd according to them is a potentially short-lived ad-hoc collection of users
bound together by some common thread – which can be communication-based,
location-based or interest based. We are inspired by this idea of formation and
detection of crowds. We try to implement detection of crowd in ``Crowdy'' in a
similar fashion.

@@

\section{Implementation}
The system is designed to run in a user's web browser.  As a result, we decided
to use some fairly standard web tools to implement the system.  The system that
we are building the user interface on top of exposes a RESTful API.  To
retrieve data, our system makes an HTTP GET request to the server.  The server
returns data in the JSON format, which is easy to decode in Javascript.

We are writing the portion of the web application that runs on the server in
python, since that is what our team has the most experience with.  We chose the
CherryPy framework to build the application because it is simple and lets us
write testable code.  We are using the Jinja templating engine to generate the
HTML pages.

>>>>

\section{Related Work}
Twitter\footnote{http://www.twitter.com} is a microblogging platform which is
fast gaining popularity\cite{Oreilly:2009} among broad sections of society and
has a global outreach spreading from developed, urban nations like the United
States where it has a high adoption rate \cite{Java:2007}, to developing
countries in parts of Asia and South America.

Along with working on understanding microblogging usage and communities
\cite{Java:2007}, the main author - Akshay Java was one of the first few who
dealt with the measurement of the usage and nature of communities in
microblogging. In his latter study \cite{Java:2008}, he presented his
observations of the microblogging phenomena and user intentions by studying the
content, topological and geographical properties of such communities. He found
that microblogging provides users with a more immediate form of communication
to talk about their daily activities and to seek or share information.

When it comes to visualization of the social crowd, a tool which in some ways
shows similar objectives as us and stands out among the rest is Vizster
\cite{Heer:2005}. It is a visualization system for which can be used for
exploration and navigation of the social networks of any scale. The system
provides facilities to improve the functionalities like discovering people or
connections which can help in promoting awareness of community structure and
exposure to information. It maintains a fun element to it while providing these
given features. It was aimed to provide specific end-users, functionalities
like situated exploration and visual analysis of big-scale communities. It is
based on a design of network layouts which have features like supporting visual
search and analysis, exploring connectivity in large graphs and automatic
identification and visualization of structures of communities.

Another related study was on how web-based retrieval from twitter can be used
as knowledge base \cite{Cheong:2009} which shows that the visualization system
we try to attempt can be helpful. Marc Cheong et al. in this study tried to
analyze demographics of the Twitter users and attempted to provide and analyze
the trend in Twitter into what's going on in the real world. They enabled us to
understand the underlying characteristics behind the trend setting in Twitter
and thereby providing a new perspective on the contributors of a trend. They
used visualization techniques along with artificial intelligence-based data
mining methods to classify the Twitter messages dealing with the trend topic.

Identifying highly dynamic ad-hoc collections of users what we refer to as
crowds in massive social messaging systems like Twitter and Facebook is
important \cite{krishna:2010}. K. Y. Kamath et al. in the study suggest an
efficient locality-based clustering approach for identifying crowds of users in
near real-time compared to more heavyweight static clustering algorithm. The
study consisted of 711,612 users and 61.3 million messages, and tells what
approaches to follow to efficiently and effectively identify Twitter-based
crowds.

In the other study based on the previous one, K. Y. Kamath et al
\cite{krishna:2011}, they add another salient feature - a novel crowd tracking
and evolution approach for linking crowds across time periods. Unlike the more
static and long-lived group-based membership offered on many social networks
their goal was to support the discovery of organic and highly-temporal group
affiliation, which they refer to as ``transient crowds''. A transient crowd
according to them is a potentially short-lived ad-hoc collection of users bound
together by some common thread – which can be communication-based,
location-based or interest based. We are inspired by this idea of formation and
detection of crowds. We try to implement detection of crowd in Crowdy in a
similar fashion.

@@

\section{Implementation}
The system is designed to run in a user's web browser.  As a result, we decided
to use some fairly standard web tools to implement the system.  The system that
we are building the user interface on top of exposes a RESTful API.  Although
it is not a part of this project, it is important to understand how the backend
works, and how it fits into this project.  First, a crawler retrieves tweets
from Twitter using Twitter's REST API, search API, and streaming API.  Next,
this data goes through several filters that add additional information to the
users and tweets.  After that, a program finds groups of users based on their
communication patterns.  This data is stored as JSON in a MongoDB database.
The backend also provides the RESTful API that we connect to.  To retrieve
data, our system makes an HTTP GET request to the server.  The server returns
data in the JSON format, which is easy to decode in Javascript.  Our GUI never
directly reads from or writes to the database.

We are writing the portion of our web application that runs on the server in
Python, since that is what our team has the most experience with.  We chose the
CherryPy framework to build the application because it is simple and lets us
write modular, testable code.  We are using the Jinja templating engine to
generate the HTML pages. This makes it easy to have a consistent look across
all of our pages without duplicating code.
